Depolarization of interacting NV$^-$ centers at low magnetic fields

### 2022-07-24 Mes papiers/Low field CR/V4/SI/Low_field_CR_SI.tex
tata

The $T_1$ profile that we observe with our samples is neither purely exponential, nor purely stretched exponential, which is to be expected when the dipole-dipole induced spin decay is comparable to the phonon induced decay.

Fig. \ref{T1_fits} shows the result of a lifetime measurement, measured with the subtraction protocol described in the main text, for zero and non-zero magnetic field. The result is fitted with exponential and stretched-exponential (with a stretch factor $\beta=0.5$) functions. We can see that for the shorter lifetime (at zero field), the measurement follows more closely the stretched-exponential profile, while the longer lifetime is closer to the exponential one (except at very short times).

In order to directly compare the results from different magnetic fields, we therefore fit each $T_1$ measurement using both a stretched and exponential decay : $S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}})$.
To add consistency to the measurement of $T_1^{\rm dd}$, we fix $T_1^{\rm ph}$ for each sample, leaving $T_1^{\rm dd}$ and the signal amplitude as the only free parameters in the fitting procedure. The value chosen for samples HPHT-150-1 and HPHT-150-2 was $T_1^{\rm ph}=3.62\ \rm ms$. Fig. 1 of the main text shows the result of fitting the same two measurement with this formula.

An other possibility would be to use an arbitrary stretched factor in the fitting function : $S(\tau)=A \exp (-(\frac{\tau}{T_1^{\rm ph}})^\beta )$. Fig. \ref{alphas} shows the optimal $\beta$ parameter as a function of the external magnetic field, which confirms that the $T_1$ profile gets closer to a pure stretched-exponential in zero magnetic field.

>>>>

%The $T_1$ profile that we observe with our samples is neither purely exponential, nor purely stretched exponential, which is to be expected when the dipole-dipole induced spin decay is comparable to the phonon induced decay.

Fig. 2-c) of the main text shows the sequence employed for measuring $T_1^{\rm dd}$. 
It consists in a pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$. In highly doped samples, this sequence often results in artifacts, mostly due to charge state transfer in the dark \citep{giri_coupled_2018, giri_selective_2019, choi_depolarization_2017}. It is therefore convenient to repeat the sequence with an additional $\pi$ pulse on one of the eight NV spin-resonances right before the spin read-out to prepare the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state \citep{jarmola_temperature-_2012,mrozek_longitudinal_2015,choi_depolarization_2017}. By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers.

Fig. \ref{T1_fits} shows the result of a lifetime measurement for the two scenarios presented in main text, namely $B=0$ and $B=50\ \rm G$ in a direction for which every classes are split, on sample HPHT-150-1. In main text, we fitted these results with the expression 
\begin{equation}
S(\tau)=A \exp(-\sqrt{\frac{\tau}{T_1^{\rm dd}}}-\frac{\tau}{T_1^{\rm ph}}),
\label{T_1 formula}
\end{equation}
in order to extract the $T_1$ part associated with the dipole-dipole coupling, as was previously done to study the cross-relaxation between NV and substitutional nitrogen (P1) center \citep{Hall}.

In Fig. \ref{T1_fits}, we show what happens when we fit the $T_1$ measurements with either purely exponential or purely stretched exponential fits. We can see that for $B=0$, which corresponds to the regime where dipole-dipole relaxation is strong, the experimental data is closely fitted by the stretched exponential profile, whereas for $B=50\ \rm G$, the experimental data follows more closely the exponential profile. In order to describe both regime simultaneously, we have to take both $T_1^{\rm dd}$ and $T_1^{\rm ph}$ into account in our analysis.

Since we are only interested in the stretched exponential part of the lifetime decay, and we are assuming that the phonon-limited exponential decay does not depend on the external magnetic field, we decided to fix the value $T_1^{\rm ph}$ for each sample, in order to reduce the number of free parameters in eq. \ref{T_1 formula}. The value chosen for all $T_1$ measurements on samples HPHT-150-1 and HPHT-150-2 was $T_1^{\rm ph}=3.62\ \rm ms$.

Another possibility to fit our $T_1$ measurements which yields satisfactory fits is to use an arbitrary stretched factor in the fitting function : 
\begin{equation}
S(\tau)=A \exp (-\left(\frac{\tau}{T_1}\right)^\beta ).
\end{equation}
 Fig. \ref{alphas} shows the optimal $\beta$ parameter as a function of the external magnetic field on sample HPHT-150-1, which confirms that the $T_1$ profile gets closer to a purely stretched exponential in zero magnetic field.
